### 2017-03-23 paper/presentation/DASFAA_presentation.tex
begin to prepare DASFAA 2017 presentation

\documentclass{beamer}

\usetheme{CambridgeUS}

\setbeamercolor{title}{bg=red!65!black,fg=white}

\setbeamertemplate{sidebar right}
{
  \vfill%
  \llap{\insertlogo\hskip0.1cm}%
  \vskip2pt%
  %\llap{\href{http://tex.stackexchange.com/}{A link to tex.sx}\hskip0.2cm}% NEW
  \vskip3pt% NEW
  \llap{\usebeamertemplate***{navigation symbols}\hskip0.1cm}%
  \vskip2pt%
}

\begin{document}

\title{Category-Level Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection}
\author{Weijing Huang, Tengjiao Wang, Wei Chen, Yazhou Wang}
\institute[Peking University]{Peking University}
\date{DASFAA 2017}
\maketitle

%------------------------------
%page 2
\begin{frame}
\frametitle{Neural Network View of Topic Models (1/2)}

The conditional probability \(p(w|d)\) is the combination of \textbf{word-topic distribution} \(p(w|t_i)\) and \textbf{topic-document distribution} \(p(t_i|d)\).
\begin{equation}
\label{eq:tm}
	p(w|d)=\sum_{i=1}^{K} p(w|t_i)p(t_i|d)
\end{equation}

Equation (\ref{eq:tm}) equals to the following form:
\begin{equation}
\label{eq:nnv_tm}
p(w|d)=\phi(w) \theta^T(d)
\end{equation}
, where \(\phi(w)\) is the row vector \textbf{word-topic distribution} \([p(w|t_1), \cdots, p(w|t_K)]\), and \(\theta(d)\) is the row vector \textbf{topic-document distribution} \([p(t_1|d), \cdots, p(t_K|d)]\).

\end{frame}

%------------------------------
%page 3
\begin{frame}
\frametitle{Neural Network View of Topic Models (2/2)}

In equation (\ref{eq:nnv_tm}), the distributions \(\phi(w)\) and \(\theta(d)\) have the following explanation:
\begin{enumerate}
	\item \(\phi(w)\) functions as the look-up layer for words with the \textbf{sigmoid} activation functions.
	\item \(\theta(d)\) functions as the look-up layer for documents with the \textbf{softmax} activation function. 
\end{enumerate}

\end{frame}

%------------------------------
%page 4
\begin{frame}
\frametitle{Neural Topic Model (NTM)}	
Symbols:

\end{frame}

\end{document}